### 2022-09-07 doc/project1.tex
Make tex files

\documentclass{article}
	\usepackage[utf8]{inputenc}
	\usepackage{mathtools}
	\usepackage{amssymb}
	\usepackage{amsmath}
	\usepackage{bm}
	\usepackage{float}
	\usepackage{listings}
	\usepackage{nicefrac}
	\usepackage{xcolor}
	\usepackage[T1]{fontenc}
	\usepackage{newtxmath,newtxtext}
	\author{Jonatan Hoffmann Hanssen \and Eric Emanuel Reber}
	\title{Project 1 FYS-STK3155}

\begin{document}
	\maketitle

\section*{Abstract}

\end{document}