### 2022-09-29 doc/project1.tex
Add beautiful proof

\documentclass{article}
	\usepackage[utf8]{inputenc}
	\usepackage{mathtools}
	\usepackage{amssymb}
	\usepackage{amsmath}
	\usepackage{bm}
	\usepackage{float}
	\usepackage{listings}
	\usepackage{nicefrac}
	\usepackage{xcolor}
	\usepackage[T1]{fontenc}
	\usepackage{newtxmath,newtxtext}
	\author{Jonatan Hoffmann Hanssen \and Eric Emanuel Reber}
	\title{Project 1 FYS-STK3155}

\begin{document}
	\maketitle

\section*{Abstract}

\section*{a}

$\bm{y}$ is defined as a $f\bm{x} + \bm{\epsilon}$. This can be approximated as a matrix multiplication $X\bm{\beta}$
plus an error vector $\bm{\epsilon}$. This means that each element of the vector
$\bm{y}$ can be expressed as follows:

\[ y_i = \sum_j x_{ij}\beta_j + \epsilon_i \]

If we take the expectation value of this expression we get the following:

\[ E[y_i] = E\left[\sum_j x_{ij}\beta_j + \epsilon_i\right] \]
\[ E[y_i] = E\left[\sum_j x_{ij}\beta_j\right] + E[\epsilon_i] \]

However, the elements of $X$ are not stochastic, and neither are the elements of $\beta$
and therefore
the first expectation value is simply the sum itself. Furthermore, $\epsilon$ is 
explicitly defined as a normal distribution $N(0,\sigma^2)$, and will by definition have
the expectation value $0$. Therefore, we end up with the final expression:

\[ E[y_i] = \sum_j x_{ij}\beta_j = \bm{X}_{i,*}\bm{\beta}\]

We can use expectation values to calculate the variance as well:

\[ var[y_i] = E\left[(y_i - E[y_i])^2\right] \]
\[ = E\left[y_i^2 - 2E[y_i]y_i + E[y_i]^2\right] \]

Distributing the outer expectation value function:

\[ var[y_i] = E[y_i^2] - 2E\left[E[y_i]\right]E[y_i] + E\left[E[y]^2\right] \]
% \[ = E[y_i^2] - 2E\left[E[y_i]\right]E[y_i] + E\left[E[y]^2\right] \]
\[ = E[y_i^2] - 2E\left[E[y_i]\right]E[y_i] + E\left[E[y]\right]E\left[E[y]\right] \]

The result of calculating the expectation value is non-stochastic. This means that
$E[E[X]] = E[X]$. From this it follows that

\[ var[y_i] = E[y_i^2] - 2E[y_i]E[y_i] + E[y_i]E[y_i] \]
\[ = E[y_i^2] - E[y_i]E[y_i] \]
\[ = E[y_i^2] - E[y_i]^2 \]

The expectation value in the second summand has been proven to be equal 
to $\bm{X}_{i,*}\bm{\beta}$ above. We therefore have

\[ var[y_i] = E[y_i^2] - (\bm{X}_{i,*}\bm{\beta})^2 \]
\[ = E\left[(X_{i,*}\bm{\beta})^2 + X_{i,*}\bm{\beta}\epsilon_i + \epsilon_i^2\right] - (\bm{X}_{i,*}\bm{\beta})^2 \]
\[ = E\left[(X_{i,*}\bm{\beta})^2\right] + E[X_{i,*}\bm{\beta}\epsilon_i] + E[\epsilon_i^2] - (\bm{X}_{i,*}\bm{\beta})^2 \]

$X_{i,*}\bm{\beta}$ and $\epsilon_i$ are both scalars. Therefore the expectation value can be written as
$E[X_{i,*}\bm{\beta}]E[\epsilon_i]$. However, $E[\epsilon_i]$ is by definition $0$, because $\epsilon_i$ is
defined as a normal distribution of mean $0$ and variance $\sigma^2$. Furthermore, the expectation value of the
non-stochastic $(X_{i,*}\bm{\beta})^2$ is simply the expression itself. We therefore have

\[ var[y_i] = (X_{i,*}\bm{\beta})^2 + E[X_{i,*}\bm{\beta}]\cdot 0 + E[\epsilon_i^2] - (\bm{X}_{i,*}\bm{\beta})^2 \]
\[ = E[\epsilon_i^2] \]

We can prove that this is equal to the variance of $\epsilon_i$:

\[ E[\epsilon_i^2] = \frac{1}{n} \sum_i \epsilon_i^2\]
\[ var[\epsilon_i] = \frac{1}{n} \sum_i (\epsilon_i - \bar{\epsilon_i})^2\]
\[ var[\epsilon_i] = \frac{1}{n} \sum_i (\epsilon_i - 0)^2\]
\[ var[\epsilon_i] = \frac{1}{n} \sum_i \epsilon_i^2 = E[\epsilon_i^2] \]

And of course, we know that the variance of $\epsilon_i$ by definition is $\sigma^2$.

\[ var[y_i] = E[\epsilon_i^2] = var[\epsilon_i] = \sigma^2 \]

\section{b}

Proof of $E[\hat{\beta}] = \beta$
$$E[\hat{\beta}] = E[(X^{T}X)^{-1}X^{T}y]$$
Because we are assuming $(X^{T}X)^{-1}X^{T}$ to be deterministic, $E[(X^{T}X)^{-1}X^{T}] = (X^{T}X)^{-1}X^{T}$. Moreover, we know $E[y] = y$, and we can thus rewrite the equation as follows:
$$E[(X^{T}X)^{-1}X^{T}y] = (X^{T}X)^{-1}X^{T}y = \beta$$
$\blacksquare$
\\

Proof of $var(\beta) = \sigma^{2}(X^{T}X)^{-1}$
$$var(\beta) = var((X^{T}X)^{-1}X^{T}y)$$
Because we are assuming $(X^{T}X)^{-1}X^{T}$ to be deterministic and $y$ to be stochastic, we can rewrite the $var(\beta)$ as following:
$$var(\beta) = (X^{T}X)^{-1}X^{T}var(y)((X^{T}X)^{-1}X^{T})^{T}$$
$$ = (X^{T}X)^{-1}X^{T}var(y)(X^{T})^{T}((X^{T}X)^{-1})^{T}$$
Since $(X^{T})^{T} = X$ and $(X^{T}X)^{-1})^{T}$ = $(X^{T}X)^{T})^{-1} = (X^{T}X)^{-1}$
$$ = (X^{T}X)^{-1}X^{T}var(y)X(X^{T}X)^{-1} $$
We know var(y) = $\sigma^{2}$, and since it is a scalar it is commutative. Thus we may move it freely
$$ = \sigma^{2}(X^{T}X)^{-1}X^{T}X(X^{T}X)^{-1} $$
$$ = \sigma^{2}(X^{T}X)^{-1} $$
$\blacksquare$

\section{c}

\subsection{Proof of the bias-variance decomposition}

Given data $\bm{y} = f(\bm{x}) + \bm{\epsilon}$ and a OLS model $\bm{\tilde{y}} = X\bm{\beta}$, we define
the Mean Squared Error as follows

\[C(X,\beta) = \frac{1}{n}\sum_{i=0}^{n-1}(y_i - \tilde{y_i})^2\]

This is simply the expectation value of $(\bm{y} - \bm{\tilde{y}})^2$

\[ C(X,\beta) = E\left[(\bm{y} - \bm{\tilde{y}})^2\right] \]

To derive the bias variance decomposition, we now add $E[\bm{\tilde{y}}] - E[\bm{\tilde{y}}]$ 
as a sneaky trick. We will also replace $\bm{y}$ with our definition $\bm{y} = \bm{f} + \bm{\epsilon}$

\[ C(X,\beta) = E\left[(\bm{f} + \bm{\epsilon} - E[\bm{\tilde{y}}] + E[\bm{\tilde{y}}] - \bm{\tilde{y}})^2\right] \]
\[ = E\left[(\bm{f} + \bm{\epsilon} - E[\bm{\tilde{y}}])^2 + 2\left[(\bm{f} + \bm{\epsilon} - E[\bm{\tilde{y}}])(E[\bm{\tilde{y}}] - \bm{\tilde{y}})\right] +(E[\bm{\tilde{y}}] - \bm{\tilde{y}})^2\right] \]
\[ = E\left[(\bm{f} + \bm{\epsilon} - E[\bm{\tilde{y}}])^2\right] + 2E\left[(\bm{f} + \bm{\epsilon} - E[\bm{\tilde{y}}])(E[\bm{\tilde{y}}] - \bm{\tilde{y}})\right] + E\left[(E[\bm{\tilde{y}}] - \bm{\tilde{y}})^2\right] \]

Let us focus our attention on the middle term first.

\[ 2E\left[(\bm{f} + \bm{\epsilon} + E[\bm{\tilde{y}}])(E[\bm{\tilde{y}}] - \bm{\tilde{y}})\right] \]

Expanding the multiplication:

\[ 2E\left[\bm{f}E[\bm{\tilde{y}}] + \bm{\epsilon}E[\bm{\tilde{y}}] - E[\bm{\tilde{y}}]E[\bm{\tilde{y}}] - \bm{f}\bm{\tilde{y}} - \bm{\epsilon}\bm{\tilde{y}} + E[\bm{\tilde{y}}]\bm{\tilde{y}}\right] \]

Distributing the expectation value:

\[ 2\left[E[\bm{f}E[\bm{\tilde{y}}]] + E[\bm{\epsilon}E[\bm{\tilde{y}}]] - E[E[\bm{\tilde{y}}]E[\bm{\tilde{y}}]] - E[\bm{f}\bm{\tilde{y}}] - E[\bm{\epsilon}\bm{\tilde{y}}] + E[E[\bm{\tilde{y}}]\bm{\tilde{y}}]\right] \]

We know that $\bm{\tilde{y}}$ and $\bm{f}$ are non-stochastic, and we know that the expectation value of $\bm{\epsilon}$ is $0$. We also know that 
$\bm{\epsilon}$ and $\bm{\tilde{y}}$ are independent, which means we can write $E[\bm{\epsilon}\bm{\tilde{y}}]$ as $E[\bm{\epsilon}]E[\bm{\tilde{y}}]$ We therefore have

\[ 2\left[\bm{f}\bm{\tilde{y}} + 0 - \bm{\tilde{y}}\bm{\tilde{y}} - \bm{f}\bm{\tilde{y}} - 0 + \bm{\tilde{y}}\bm{\tilde{y}}\right] = 0 \]

We then have

\[ C(X,\beta) = E\left[(\bm{f} + \bm{\epsilon} - E[\bm{\tilde{y}}])^2\right] + E\left[(E[\bm{\tilde{y}}] - \bm{\tilde{y}})^2\right] \]

The second term is by definition the variance of model.

\[ C(X,\beta) = E\left[(\bm{f} + \bm{\epsilon} - E[\bm{\tilde{y}}])^2\right] + var[\bm{\tilde{f}}] \]

And now the final term

\[ E\left[(\bm{f} + \bm{\epsilon} - E[\bm{\tilde{y}}])^2\right] \]

\[ = E\left[\bm{f}^2 + \bm{f}\bm{\epsilon} - \bm{f}E[\bm{\tilde{y}}] + \bm{f\epsilon} + \bm{\epsilon}^2 - \bm{\epsilon}E[\bm{\tilde{y}}] - \bm{f}E[\bm{\tilde{y}}] - \bm{\epsilon}E[\bm{\tilde{y}}] + E[\bm{\tilde{y}}]^2 \right] \]

After distributing the expectation value we get the following, keeping in mind that $E[\bm{f}] = E[\bm{y} - \bm{\epsilon}] = E[\bm{y}] - E[\bm{\epsilon}] = \bm{y}$

\[ \bm{y}^2 - 2\bm{y}\bm{\tilde{y}} + E[\bm{\tilde{y}}]^2 + \sigma^2\]

\[ (\bm{y} + E[\bm{\tilde{y}}]^2 + \sigma^2 \]

We see that the quadratic form is equal to $bias[\bm{\tilde{y}}]^2$. We can then conclude that the cost function can be rewritten as follows:

\[ C(X,\beta) = bias[\bm{\tilde{y}}]^2 + var[\bm{\tilde{f}}] + \sigma^2 \]

% The first term here represents the bias, the second term is equal to zero and the last term represents the variance.

\end{document}